\documentclass[journal]{IEEEtai}

\input{packages}
\graphicspath{{./plots/}}
\input{macros}

\begin{document}

%% \jvol{XX}
%% \jnum{XX}
%% \paper{1234567}
%% \pubyear{2020}
%% \publisheddate{xxxx 00, 0000}
%% \currentdate{xxxx 00, 0000}
%% \doiinfo{TQE.2020.Doi Number}
\title{Supervised Learning for Table Tennis Match Prediction}
\author{\IEEEauthorblockN{Yu-hung Sophie Chiang}, and
    Gyorgy Denes\,\orcidlink{0000-0003-4792-9100}

\thanks{Manuscript submitted for review on  June 2nd 2022} 
\thanks{Yu-hung Sophie Chiang and Gyorgy Denes are with the Perse School, Hills Rd, Cambridge CB2 8QF, UK (correspondence email: gdenes@perse.co.uk)}
}

\markboth{Journal of IEEE Transactions on Artificial Intelligence, Vol. 00, No. 0, Month 2022}
{Chiang and Denes: Supervised Learning for Table Tennis Match Prediction}



\maketitle

%\thispagestyle{plain}
%\pagestyle{plain}

\begin{abstract}
Machine learning, classification and prediction models have applications across a range of fields. Sport analytics is one such increasingly popular field, but most existing work focuses on automated refereeing and injury prevention in mainstream sports. 
Research on other sports, such as table tennis, have only recently started gaining more traction. In this paper, we aim to predict the outcome of table tennis matches; we valuate a range of existing supervised machine learning models to find a reliable predictor, trained on historical player and  match statistics. We also derive 12 features and demonstrate their utility in an ablation study. We found that  logistic regression performed best on the test data, although differences between models were not always significant.  Our results match the accuracy of state-of-the-art in comparable sports, such as tennis.
\end{abstract}

\begin{IEEEImpStatement}
The use of supervised learning models has been proposed for outcome prediction in a number of sports; however, there has been little research done for table tennis. This paper extends the application field in this direction, rigorously evaluating existing models. Furthermore, we also present a set of additional features that increase model accuracy, as validated in the ablation study. The results can serve as a baseline for future table tennis prediction models, and can feed back to prediction research in similar ball sports.
\end{IEEEImpStatement}


\if{false}  % Feel free to add here, but removing from manuscript
If not IEEE TAI, we could consider:
\begin{itemize}
    \item{Computers \& Mathematics with Applications (Elsevier)}
    \item{\sophie{MLSA 2022 : Machine Learning and Data Mining for Sports Analytics}}
\end{itemize}
\fi

% GXD: roughly the structure I was thinking, but feel free to tweak/move bits
\input{tex/introduction}
\input{tex/related}
\input{tex/dataset}
\input{tex/features}
\input{tex/models}
\input{tex/results}
\input{tex/discussion}
\input{tex/conclusions}

\section*{Acknowledgements}
The authors would also like to thank the OSAI team for granting permission to use their dataset.

\bibliographystyle{ieeetr}
\bibliography{References}
\end{document}
